How to select predictive models for decision-making or causal inference?

### 2024-06-02 how_to_select.tex
gigascience: new jounrnal compiling format

% XXX: for submission
%\documentclass[10pt,twocolumn]{article}
\documentclass[11pt]{article}
%\documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%

%\usepackage[a4paper, left=15mm, right=15mm, top=20mm,]{geometry}
% XXX: for submission
\usepackage[a4paper, left=25mm, right=25mm, top=20mm,]{geometry}
% declared functions and additional packages
\usepackage{caption}
\usepackage{subcaption}
\usepackage{threeparttable}
\usepackage{times}
\usepackage{hyperref}
\def\mycitecolor{green!50!black}
\def\mylinkcolor{red!60!black}
\hypersetup{ colorlinks=true, linkcolor=\mylinkcolor, citecolor=\mycitecolor }
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb} %mathbb packages
\usepackage{siunitx}
\usepackage{algorithm, algorithmicx, algpseudocode}
\usepackage{multirow}
\usepackage{appendix}
\usepackage[sorting=nyt,style=numeric,natbib=true,doi=false,isbn=false,url=false,eprint=false, maxnames=999, maxcitenames=2]{biblatex} 
% To cite with superscript
\let\cite=\supercite
\DeclareNameAlias{default}{last-first}
\usepackage{placeins}
% associated arxiv class
%\usepackage{arxiv}
\section{Methods}\label{sec:framework}
\subsection{Neyman-Rubin Potential Outcomes framework}%
\label{sec:neyman_rubin}%

\begin{table*}[bt!]
    \makebox[\textwidth]{
        \begin{threeparttable}[b]
            \caption{Review of causal risks
                ---
                The $R\text{-risk}^*$ is called $\tau \text{-risk}_R$ in
                \citet{schuler_comparison_2018}.
                \label{tab:evaluation_metrics}}
            \centering
            \begin{tabular}{llr}
                \toprule
                Risk                                                                                           & Equation
                                                                                                               & Reference                                                                                                                                           \\
                \midrule
                $\tau\text{-risk}=\text{MSE}(\tau(X), \tau_f(X))$                                              & $\mathbb E_{X\sim
                            p(X)}[(\tau(X) - \hat \tau_f(X))^2] $
                                                                                                               & Eq. \ref{eq:tau_risk} \cite{hill_bayesian_2011}                                                                                                     \\
                $\mu\text{-risk} = \text{MSE}(Y, f(X))$                                                        & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D}\left[(Y-f(X ; A))^2 \right]$
                                                                                                               & Def. \ref{def:mu_risk} \cite{schuler_comparison_2018}                                                                                               \\
                $\mu\text{-risk}_{IPW}^*$                                                                      & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D}\left[ \Big( \frac{A}{e(X)} + \frac{1-A}{1-e(X)} \Big)
                (Y-f(X ; A))^2 \right]$                                                                        & Def.
                \ref{def:mu_ipw_risk} \cite{vanderlaan_unified_2003}                                                                                                                                                                                                 \\
                $\tau\text{-risk}^{\star}_{IPW}$                                                               & $\mathbb{E}_{(Y, X, A) \sim \mathcal D} \left[ \Big(Y \left( \frac{A}{e(X)} - \frac{1-A}{1-e(X)}\right)-\hat \tau_f\left(X\right)\Big)^2 \right]$ &
                Def. \ref{def:tau_ipw_risk} \cite{wager_estimation_2018}
                \\
                $U\text{-risk}^*$                                                                              & $\mathbb{E}_{(Y, X, A) \sim \mathcal D}  \big[
                \big( \frac{Y-m\left(X\right)}{A-e\left(X\right)} -  \hat \tau_f\left(X\right)\big)^{2} \big]$ &
                Def. \ref{def:u_risk} \cite{nie_quasioracle_2017}
                \\
                $R\text{-risk}^*$                                                                              & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D} \big[\big(\left(Y-m\left(X\right)\right)
                -\left(A-e\left(X\right)\right) \hat \tau_f\left(X\right)\big)^{2} \big]$                      &
                Def. \ref{def:r_risk} \cite{nie_quasioracle_2017}
                \\
                \bottomrule
            \end{tabular}
        \end{threeparttable}
    }
\end{table*}

\begin{figure}[h!]
    % XXX: for submission
    \centering\begin{minipage}{.6\linewidth}
        \includegraphics[width=\linewidth]{images/estimation_procedure_causal_selection_procedure.png}
    \end{minipage}
    \caption{Estimation procedure for causal model
        selection.}\label{problem:estimation_procedure:figure}
\end{figure}

\setlength{\columnsep}{0.5cm}
\usepackage{wrapfig}
%\usepackage{wraptable}
\usepackage{authblk}% for author affiliations
\usepackage[dvipsnames]{xcolor}
\usepackage{setspace} % double spaced required by journal
\doublespacing

\usepackage[%
    font={footnotesize,sf},
    labelfont=bf,
    format=hang,    
    format=plain,
    margin=0pt,
]{caption}

% TODO commands
\usepackage{snaptodo}
\setlength{\marginparwidth}{2.3cm}
\setlength{\marginparsep}{.1cm}
\snaptodoset{block rise=2em}
\snaptodoset{margin block/.style={font=\tiny}}

\newcommand{\gv}[1]{\snaptodo[margin block/.append style=green!50!black]{\sloppy\textbf{GV}: #1}}

% Uncomment below to erase all comments
%\renewcommand{\snaptodo}[2][1]{\relax}

% XXX: for submission
\usepackage{setspace}

\usepackage{comment}
\setcounter{secnumdepth}{0}

%\setlength{\marginparwidth}{0.5cm}
%\setlength\textwidth{177.8mm}

% maths commands
\newcommand{\indep}{\perp \!\!\! \perp}
\newcommand{\eg}{\emph{e.g.}}
\DeclareMathOperator*{\argmin}{argmin} \def\mycitecolor{green!50!black}
\usepackage{mathtools}
\newcommand\myeq{\stackrel{\mathclap{\text{def}}}{=}}

\newtheorem{definition}{Definition}
\newtheorem{condition}{Condition}
\newtheorem{property}{Property}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}
\newtheorem{proposition*}{Proposition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{assumption}{Assumption}
\newtheorem{proof}{Proof}
%
\def\keywordname{{\bfseries \emph{Keywords}}}%
\def\keywords#1{\par\addvspace\medskipamount{\rightskip=0pt plus1cm
\def\and{\ifhmode\unskip\nobreak\fi\ $\cdot$
}\noindent\keywordname\enspace\ignorespaces#1\par}}

\title{How to select predictive models for decision-making or causal inference?}
\addbibresource{biblio.bib}

\author[1,2,*]{Matthieu Doutreligne, PhD}
\author[1]{Gaël Varoquaux, PhD}

\affil[1]{Inria, Soda, Saclay, France}
\affil[2]{Mission Data, Haute Autorité de Santé, Saint-Denis, France}
\affil[*]{Corresponding author: matt.dout@gmail.com}

%\raggedbottom

@@

%\corres{*\email{matthieu.doutreligne@inria.fr}}

\onecolumn
\maketitle
\begin{abstract}\noindent
    \textbf{Background:}
    We investigate which procedure selects the predictive model most
    trustworthy to reason on the effect of an intervention and support
    decision making.\\
    %As predictive models --\eg from machine learning-- give likely
    %outcomes, they may be used to, a
    %causal-inference task. The increasing complexity of health data has opened
    %the door to a plethora of models, but also the Pandora box of model
    %selection: which of these models yield the most valid causal estimates? 
    %\textbf{Methods:}
    We study such a variety of model selection procedures in practical settings:
    finite samples settings and without theoretical assumption of
    \emph{well-specified} models. Beyond standard cross-validation or
    internal validation procedures, we also study elaborate causal risks.
    These build proxies of the causal error using ``nuisance''
    re-weighting to compute it on the observed data. We evaluate whether
    empirically estimated nuisances, which are necessarily noisy,
    add noise to model selection. We compare different metrics for
    causal model selection in an extensive empirical
    study based on a simulation and three healthcare datasets
    based on real covariates. \\
    \textbf{Results:} Among all metrics, the mean squared error,
    classically used to evaluate predictive modes,
    is worse. Re-weighting it with propensity score
    does not bring much improvements. The
    $R\text{-risk}$, which uses as nuisances a model of
    mean outcome and propensity scores, leads to the best performances.
    Nuisance corrections are best estimated with flexible estimators such
    as a super learner.
    \\
    \textbf{Conclusions:} When predictive models are used to
    reason on the effect of an intervention, they must be evaluated
    with different procedures than standard predictive
    settings; using the $R\text{-risk}$ from causal inference.
\end{abstract}
\keywords{Model Selection, Predictive model, Treatment Effect, G-formula, Machine Learning, }
\clearpage

% XXX: for submission
%\twocolumn
% XXX: for submission (double spaced required)
%\setstretch{1.1}
\doublespacing
\section{Introduction}\label{sec:intro}

@@

        \setlength{\unitlength}{.1\linewidth}
        \begin{picture}(10,5.2)(0,0)
            \put(0, 0){\includegraphics[width=1\linewidth]{images/toy_random_forest_high_R2_high_tau_risk.pdf}}%
            \put(7.05,4.4){\vector(0,1){.47}}
            \put(7.05,4.4){\vector(0,-1){.55}}
            \put(7.1,4.5){\scriptsize\sffamily treatment effect}
            \put(7.1,4.15){\small $\tau(x)$}
            \put(5.2,3.6){\small $\hat{\tau}(x)$}
            \put(5.15,3.6){\vector(0,1){.29}}
            \put(5.15,3.6){\vector(0,-1){.15}}
        \end{picture}

@@

        \hfill%
        \includegraphics[width=1\linewidth]{images/toy_tlinear_model_small_R2_small_tau_risk.pdf}%
    \end{minipage}

@@

\begin{table*}[bt!]
    \makebox[\textwidth]{
        \begin{threeparttable}[b]
            \caption{Review of causal risks
                ---
                The $R\text{-risk}^*$ is called $\tau \text{-risk}_R$ in
                \citet{schuler_comparison_2018}.
                \label{tab:evaluation_metrics}}
            \centering
            \begin{tabular}{llr}
                \toprule
                Risk                                                                                           & Equation
                                                                                                               & Reference                                                                                                                                           \\
                \midrule
                $\tau\text{-risk}=\text{MSE}(\tau(X), \tau_f(X))$                                              & $\mathbb E_{X\sim
                            p(X)}[(\tau(X) - \hat \tau_f(X))^2] $
                                                                                                               & Eq. \ref{eq:tau_risk} \cite{hill_bayesian_2011}                                                                                                     \\
                $\mu\text{-risk} = \text{MSE}(Y, f(X))$                                                        & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D}\left[(Y-f(X ; A))^2 \right]$
                                                                                                               & Def. \ref{def:mu_risk} \cite{schuler_comparison_2018}                                                                                               \\
                $\mu\text{-risk}_{IPW}^*$                                                                      & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D}\left[ \Big( \frac{A}{e(X)} + \frac{1-A}{1-e(X)} \Big)
                (Y-f(X ; A))^2 \right]$                                                                        & Def.
                \ref{def:mu_ipw_risk} \cite{vanderlaan_unified_2003}                                                                                                                                                                                                 \\
                $\tau\text{-risk}^{\star}_{IPW}$                                                               & $\mathbb{E}_{(Y, X, A) \sim \mathcal D} \left[ \Big(Y \left( \frac{A}{e(X)} - \frac{1-A}{1-e(X)}\right)-\hat \tau_f\left(X\right)\Big)^2 \right]$ &
                Def. \ref{def:tau_ipw_risk} \cite{wager_estimation_2018}
                \\
                $U\text{-risk}^*$                                                                              & $\mathbb{E}_{(Y, X, A) \sim \mathcal D}  \big[
                \big( \frac{Y-m\left(X\right)}{A-e\left(X\right)} -  \hat \tau_f\left(X\right)\big)^{2} \big]$ &
                Def. \ref{def:u_risk} \cite{nie_quasioracle_2017}
                \\
                $R\text{-risk}^*$                                                                              & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D} \big[\big(\left(Y-m\left(X\right)\right)
                -\left(A-e\left(X\right)\right) \hat \tau_f\left(X\right)\big)^{2} \big]$                      &
                Def. \ref{def:r_risk} \cite{nie_quasioracle_2017}
                \\
                \bottomrule
            \end{tabular}
        \end{threeparttable}
    }
\end{table*}

@@

\begin{figure}[h!]
    % XXX: for submission
    \centering\begin{minipage}{.6\linewidth}
        \includegraphics[width=\linewidth]{images/estimation_procedure_causal_selection_procedure.png}
    \end{minipage}
    \caption{Estimation procedure for causal model
        selection.}\label{problem:estimation_procedure:figure}
\end{figure}

@@

\begin{figure}[!b]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{images/_1_r_risk_domination_r_risk_domination__ref_metric_mean_risks_by_Dataset_feasible_only.pdf}
    \end{minipage}
    \caption{\textbf{The $R$-risk is the best metric}: Relative Kendall's $\tau$ agreement with $\tau\text{-risk}$.
        Strong and Weak overlap correspond to the first and last tertiles of the overlap distribution measured with
        Normalized Total Variation eq. \ref{eq:ntv}. \ref{apd:experiments:additional_results} presents the same results
        by adding semi-oracle risks in Figure \ref{apd:fig:relative_kendalls_all_datasets_all_metrics}, measured with
        absolute Kendall's in Figure \ref{apd:fig:all_datasets_tau_risk_ranking_agreement} and with $\tau\mathrm{-risk}$
        gains in Figure \ref{apd:all_datasets_normalized_bias_tau_risk_to_best_method}. Table
        \ref{apd:table:relative_kendalls_all_datasets} gives median and
        IQR of the relative Kendall.}\label{fig:relative_kendalls_all_datasets}
\end{figure}

@@

\begin{figure}[!h]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{images/_2_overlap_influence_overlap_by_bin_comparaison_kendall_by_Dataset_r_risk_only_twocols.pdf}
    \end{minipage}
    \caption{\textbf{Model selection is harder for low population
            overlap}:
        Kendall's $\tau$ agreement with $\tau\text{-risk}$. Strong, medium and Weak overlap
        are the tertiles of the overlap measured with NTV eq. \ref{eq:ntv}. \ref{apd:experiments:additional_results} presents results for all
        metrics in Figure \ref{apd:fig:all_datasets_overlap_effect} in absolute
        Kendall's and continuous overlap values in Figure
            {\ref{apd:fig:all_datasets_tau_risk_ranking_agreement}}.}\label{fig:all_datasets_overlap_effect_r_risk}
\end{figure}

@@

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{images/_3_procedure_r_risk_only_3datasets_twocols.pdf}
    \end{minipage}

@@

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{images/_4_nuisance_models_r_risk_only_3datasets_twocols.pdf}
    \end{minipage}

@@

\section*{Acknowledgments}

\subsection*{Financial disclosure}

None reported.

\subsection*{Conflict of interest}

The authors declare no potential conflict of interests.

%\if@Mybibrefstyle%
\printbibliography
%\fi if using njdnatbib.sty \bibliography{bibliography}%

@@

@@

    \part*{Supplementary materials:\\
        How to select predictive models for decision-making or causal inference? }

@@

    \begin{figure}[!b]
        \centering
        \includegraphics[width=0.95\linewidth]{images/2023-03-08-11-10-28_acic_2016_ate_heterogeneity.parquet_abs_bias_ylog_scale=True.pdf}%
        \caption{\textbf{Different outcome models lead to different
                estimation errors on the Average Treatment Effects},
            on 77 classic simulations with known true causal effect
            \cite{dorie_automated_2019}. The different models are ridge regression
            and random forests with different hyper-parameters
            (details
            \ref{apd:toy_example:acic_2016_ate_variability}). The different configurations are
            plotted as a function of increasing difference between treated and
            untreated population --see
            \autoref{subsec:measuring_overlap}.
            There is no systematic best performer; data-driven model
            selection is important.
            \label{fig:acic_2016_ate_heterogeneity}%
        }
    \end{figure}

@@

    Using eight simulations setups from \cite{powers_methods_2018}, where
    the oracle CATE is known, \citet{schuler_comparison_2018} compare four
    causal risks, concluding that for CATE estimation the best
    model-selection risk is the so-called $R\text{-risk}$
    \cite{nie_quasioracle_2017} --def.\,\ref{def:r_risk}, below. Their
    empirical results are clear for randomized treatment allocation but less
    convincing for observational settings where both simple Mean Squared
    Error --MSE, $\mu\text{-risk}(f)$ def.\,\ref{def:mu_risk}-- and
    reweighted MSE --$\mu\text{-risk}_{IPW}$ def.\,\ref{def:mu_ipw_risk}--
    appear to perform better than $R\text{-risk}$ on half of the simulations.
    Another work \cite{alaa_validating_2019} studied empirically both MSE and
    reweighted MSE risks on the semi-synthetic ACIC 2016 datasets
    \cite{dorie_automated_2019}, but did not include the $R\text{-risk}$. We complete these
    prior empirical work by studying a wider variety of data generative
    processes and varying the influence of overlap, an important parameter of
    the data generation process which makes a given causal metric appropriate
    \cite{damour_overlap_2020}. We also study how to best adapt
    cross-validation procedures to causal metrics which themselves come with
    models to estimate.

@@

    By contrast, \citet{johansson2022generalization} studies causal estimation
    without assuming that estimators are well specified. They derive an upper bound
    on the oracle error to the CATE ($\tau\text{-risk}$) that involves the error on
    the outcome and the similarity of the distributions of treated and control
    patients. However, they use this upper bound for model optimization,
    and do not give insights on model selection. In addition, for hyperparameter
    selection, they rely on a plugin estimate of the $\tau\text{-risk}$ built with
    counterfactual nearest neighbors, which has been shown ineffective
    \cite{schuler_comparison_2018}.

@@

    \begin{figure}[!t]
        \centering
        \includegraphics[width=\linewidth]{images/caussim_example_rs_gaussian=8_rs_rotation=8_ntv=0.37_D=2_overlap=0.7_p_A=0.1.pdf}

>>>>

%%
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

%%%CONTEMPORARY%%%
\documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
%\documentclass[unnumsec,webpdf,contemporary,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,contemporary,small]{oup-authoring-template}

%%%MODERN%%%
%\documentclass[unnumsec,webpdf,modern,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,modern,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,small]{oup-authoring-template}

%%%TRADITIONAL%%%
%\documentclass[unnumsec,webpdf,traditional,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,namedate,webpdf,traditional,medium]{oup-authoring-template}
%\documentclass[namedate,webpdf,traditional,small]{oup-authoring-template}

%\onecolumn % for one column layouts

%\usepackage{showframe}

\graphicspath{{images/}}

% line numbers
%\usepackage[mathlines, switch]{lineno}
%\usepackage[right]{lineno}

\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}

% added commands and packages
\newtheorem{assumption}{Assumption}
\newtheorem{proposition*}{Proposition}
\newcommand{\indep}{\perp \!\!\! \perp}
\newcommand{\eg}{\emph{e.g.}}
\DeclareMathOperator*{\argmin}{argmin} \def\mycitecolor{green!50!black}
\usepackage{mathtools}
\newcommand\myeq{\stackrel{\mathclap{\text{def}}}{=}}
\usepackage{booktabs} % For formal tables (eg. bottomrules command)
\usepackage{threeparttable} % For table 
\usepackage{caption} % for subfigures
\usepackage{subcaption}
\usepackage{placeins} % for FloatBarrier command to prevent a figure to pass some point.

\bibliographystyle{unsrtnat}
% add external bibliography
%\usepackage[sorting=nyt,style=numeric,natbib=true,doi=false,isbn=false,url=false,eprint=false, maxnames=999, maxcitenames=2]{biblatex} 
%\let\cite=\supercite
%\DeclareNameAlias{default}{last-first}
%\addbibresource{biblio.bib}

@@

\journaltitle{GigaScience}
\DOI{DOI HERE}
\copyrightyear{2024}
\pubyear{2024}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}

\firstpage{1}

%\subtitle{Subject Section}

\title[How to select predictive models for decision-making or causal inference?]{How to select predictive models for decision-making or causal inference?}

\author[1, 2 $\ast$]{Matthieu Doutreligne}
\author[1]{Gaël Varoquaux}

\authormark{Author Name et al.}

\address[1]{\orgdiv{Soda}, \orgname{Inria, Saclay}, \orgaddress{\country{France}}}
\address[2]{\orgdiv{Mission Data}, \orgname{Haute Autorité de Santé}, \orgaddress{\country{France}}}

\corresp[$\ast$]{Corresponding author. \href{email:matt.dout@gmail.com}{matt.dout@gmail.com}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

%\editor{Associate Editor: Name}

%\abstract{
%\textbf{Motivation:} .\\
%\textbf{Results:} .\\
%\textbf{Availability:} .\\
%\textbf{Contact:} \href{name@email.com}{name@email.com}\\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Journal Name}
%online.}

\abstract{\textbf{Background:}
    We investigate which procedure selects the predictive model most
    trustworthy to reason on the effect of an intervention and support
    decision making.\\
    %As predictive models --\eg from machine learning-- give likely
    %outcomes, they may be used to, a
    %causal-inference task. The increasing complexity of health data has opened
    %the door to a plethora of models, but also the Pandora box of model
    %selection: which of these models yield the most valid causal estimates? 
    %\textbf{Methods:}
    We study such a variety of model selection procedures in practical settings:
    finite samples settings and without theoretical assumption of
    \emph{well-specified} models. Beyond standard cross-validation or
    internal validation procedures, we also study elaborate causal risks.
    These build proxies of the causal error using ``nuisance''
    re-weighting to compute it on the observed data. We evaluate whether
    empirically estimated nuisances, which are necessarily noisy,
    add noise to model selection. We compare different metrics for
    causal model selection in an extensive empirical
    study based on a simulation and three healthcare datasets
    based on real covariates. \\
    \textbf{Results:} Among all metrics, the mean squared error,
    classically used to evaluate predictive modes,
    is worse. Re-weighting it with propensity score
    does not bring much improvements. The
    $R\text{-risk}$, which uses as nuisances a model of
    mean outcome and propensity scores, leads to the best performances.
    Nuisance corrections are best estimated with flexible estimators such
    as a super learner.
    \\
    \textbf{Conclusions:} When predictive models are used to
    reason on the effect of an intervention, they must be evaluated
    with different procedures than standard predictive
    settings; using the $R\text{-risk}$ from causal inference.}
\keywords{Model Selection, Predictive model, Treatment Effect, G-formula, Machine Learning, }

% \boxedtext{
% \begin{itemize}
% \item Key boxed text here.
% \item Key boxed text here.
% \item Key boxed text here.
% \end{itemize}}

\maketitle

\section{Introduction}

@@

        \setlength{\unitlength}{.1\linewidth}
        \begin{picture}(10,5.2)(0,0)
            \put(0, 0){\includegraphics[width=1\linewidth]{toy_random_forest_high_R2_high_tau_risk.pdf}}%
            \put(7.05,4.4){\vector(0,1){.47}}
            \put(7.05,4.4){\vector(0,-1){.55}}
            \put(7.1,4.5){\scriptsize\sffamily treatment effect}
            \put(7.1,4.15){\small $\tau(x)$}
            \put(5.2,3.6){\small $\hat{\tau}(x)$}
            \put(5.15,3.6){\vector(0,1){.29}}
            \put(5.15,3.6){\vector(0,-1){.15}}
        \end{picture}

@@

        \hfill%
        \includegraphics[width=1\linewidth]{toy_tlinear_model_small_R2_small_tau_risk.pdf}%
    \end{minipage}

@@

\begin{table*}[bt!]
    \makebox[\textwidth]{
        \begin{threeparttable}[b]
            \caption{Review of causal risks
                ---
                The $R\text{-risk}^*$ is called $\tau \text{-risk}_R$ in
                \cite{schuler_comparison_2018}.
                \label{tab:evaluation_metrics}}
            \centering
            \begin{tabular}{llr}
                \toprule
                Risk                                                                                           & Equation
                                                                                                               & Reference                                                                                                                                           \\
                \midrule
                $\tau\text{-risk}=\text{MSE}(\tau(X), \tau_f(X))$                                              & $\mathbb E_{X\sim
                            p(X)}[(\tau(X) - \hat \tau_f(X))^2] $
                                                                                                               & Eq. \ref{eq:tau_risk} \cite{hill_bayesian_2011}                                                                                                     \\
                $\mu\text{-risk} = \text{MSE}(Y, f(X))$                                                        & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D}\left[(Y-f(X ; A))^2 \right]$
                                                                                                               & Def. \ref{def:mu_risk} \cite{schuler_comparison_2018}                                                                                               \\
                $\mu\text{-risk}_{IPW}^*$                                                                      & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D}\left[ \Big( \frac{A}{e(X)} + \frac{1-A}{1-e(X)} \Big)
                (Y-f(X ; A))^2 \right]$                                                                        & Def.
                \ref{def:mu_ipw_risk} \cite{vanderlaan_unified_2003}                                                                                                                                                                                                 \\
                $\tau\text{-risk}^{\star}_{IPW}$                                                               & $\mathbb{E}_{(Y, X, A) \sim \mathcal D} \left[ \Big(Y \left( \frac{A}{e(X)} - \frac{1-A}{1-e(X)}\right)-\hat \tau_f\left(X\right)\Big)^2 \right]$ &
                Def. \ref{def:tau_ipw_risk} \cite{wager_estimation_2018}
                \\
                $U\text{-risk}^*$                                                                              & $\mathbb{E}_{(Y, X, A) \sim \mathcal D}  \big[
                \big( \frac{Y-m\left(X\right)}{A-e\left(X\right)} -  \hat \tau_f\left(X\right)\big)^{2} \big]$ &
                Def. \ref{def:u_risk} \cite{nie_quasioracle_2017}
                \\
                $R\text{-risk}^*$                                                                              & $\mathbb{E}_{(Y, X, A)
                        \sim \mathcal D} \big[\big(\left(Y-m\left(X\right)\right)
                -\left(A-e\left(X\right)\right) \hat \tau_f\left(X\right)\big)^{2} \big]$                      &
                Def. \ref{def:r_risk} \cite{nie_quasioracle_2017}
                \\
                \bottomrule
            \end{tabular}
        \end{threeparttable}
    }
\end{table*}

@@

\begin{figure}[h!]

    \centering\begin{minipage}{.6\linewidth}
        \includegraphics[width=\linewidth]{estimation_procedure_causal_selection_procedure.png}
    \end{minipage}
    \caption{Estimation procedure for causal model
        selection.}\label{problem:estimation_procedure:figure}
\end{figure}

@@

\begin{figure}[!b]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{_1_r_risk_domination_r_risk_domination__ref_metric_mean_risks_by_Dataset_feasible_only.pdf}
    \end{minipage}
    \caption{\textbf{The $R$-risk is the best metric}: Relative Kendall's $\tau$ agreement with $\tau\text{-risk}$.
        Strong and Weak overlap correspond to the first and last tertiles of the overlap distribution measured with
        Normalized Total Variation eq. \ref{eq:ntv}. \ref{apd:experiments:additional_results} presents the same results
        by adding semi-oracle risks in Figure \ref{apd:fig:relative_kendalls_all_datasets_all_metrics}, measured with
        absolute Kendall's in Figure \ref{apd:fig:all_datasets_tau_risk_ranking_agreement} and with $\tau\mathrm{-risk}$
        gains in Figure \ref{apd:all_datasets_normalized_bias_tau_risk_to_best_method}. Table
        \ref{apd:table:relative_kendalls_all_datasets} gives median and
        IQR of the relative Kendall.}\label{fig:relative_kendalls_all_datasets}
\end{figure}

@@

\begin{figure}[!h]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{_2_overlap_influence_overlap_by_bin_comparaison_kendall_by_Dataset_r_risk_only_twocols.pdf}
    \end{minipage}
    \caption{\textbf{Model selection is harder for low population
            overlap}:
        Kendall's $\tau$ agreement with $\tau\text{-risk}$. Strong, medium and Weak overlap
        are the tertiles of the overlap measured with NTV eq. \ref{eq:ntv}. \ref{apd:experiments:additional_results} presents results for all
        metrics in Figure \ref{apd:fig:all_datasets_overlap_effect} in absolute
        Kendall's and continuous overlap values in Figure
            {\ref{apd:fig:all_datasets_tau_risk_ranking_agreement}}.}\label{fig:all_datasets_overlap_effect_r_risk}
\end{figure}

@@

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{_3_procedure_r_risk_only_3datasets_twocols.pdf}
    \end{minipage}

@@

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{.75\linewidth}
        \includegraphics[width=\linewidth]{_4_nuisance_models_r_risk_only_3datasets_twocols.pdf}
    \end{minipage}

@@

@@

%%%%%%%%%%%%%%

@@

    %\part*{Supplementary materials:\\How to select predictive models for decision-making or causal inference? }

@@

    \begin{figure}[!b]
        \centering
        \includegraphics[width=0.95\linewidth]{2023-03-08-11-10-28_acic_2016_ate_heterogeneity.parquet_abs_bias_ylog_scale=True.pdf}%
        \caption{\textbf{Different outcome models lead to different
                estimation errors on the Average Treatment Effects},
            on 77 classic simulations with known true causal effect
            \cite{dorie_automated_2019}. The different models are ridge regression
            and random forests with different hyper-parameters
            (details
            \ref{apd:toy_example:acic_2016_ate_variability}). The different configurations are
            plotted as a function of increasing difference between treated and
            untreated population --see
            \autoref{subsec:measuring_overlap}.
            There is no systematic best performer; data-driven model
            selection is important.
            \label{fig:acic_2016_ate_heterogeneity}%
        }
    \end{figure}

@@

    Using eight simulations setups from \cite{powers_methods_2018}, where
    the oracle CATE is known, \cite{schuler_comparison_2018} compare four
    causal risks, concluding that for CATE estimation the best
    model-selection risk is the so-called $R\text{-risk}$
    \cite{nie_quasioracle_2017} --def.\,\ref{def:r_risk}, below. Their
    empirical results are clear for randomized treatment allocation but less
    convincing for observational settings where both simple Mean Squared
    Error --MSE, $\mu\text{-risk}(f)$ def.\,\ref{def:mu_risk}-- and
    reweighted MSE --$\mu\text{-risk}_{IPW}$ def.\,\ref{def:mu_ipw_risk}--
    appear to perform better than $R\text{-risk}$ on half of the simulations.
    Another work \cite{alaa_validating_2019} studied empirically both MSE and
    reweighted MSE risks on the semi-synthetic ACIC 2016 datasets
    \cite{dorie_automated_2019}, but did not include the $R\text{-risk}$. We complete these
    prior empirical work by studying a wider variety of data generative
    processes and varying the influence of overlap, an important parameter of
    the data generation process which makes a given causal metric appropriate
    \cite{damour_overlap_2020}. We also study how to best adapt
    cross-validation procedures to causal metrics which themselves come with
    models to estimate.

@@

    By contrast, without assuming that estimators are well specified, there exists an upper bound
    on the oracle error to the CATE ($\tau\text{-risk}$) that involves the error on
    the outcome and the similarity of the distributions of treated and control
    patients \cite{johansson2022generalization}. However, they use this upper bound for model optimization,
    and do not give insights on model selection. In addition, for hyperparameter
    selection, they rely on a plugin estimate of the $\tau\text{-risk}$ built with
    counterfactual nearest neighbors, which has been shown ineffective
    \cite{schuler_comparison_2018}.

@@

    \begin{figure}[!t]
        \centering
        \includegraphics[width=\linewidth]{caussim_example_rs_gaussian=8_rs_rotation=8_ntv=0.37_D=2_overlap=0.7_p_A=0.1.pdf}
